---
title: "BRCA Somatic Recurrent in 20 Samples"
author: "C.J. Liu"
date: "11/18/2016"
output:
  word_document:
    toc: yes
  html_document:
    depth: 3
    highlight: tango
    number_sections: yes
    theme: united
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load library}
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
library(knitr)
library(biomaRt)
```
## Samples
```{r samples}
sampleHeaders = c("sample.type","file.name","file.id","submitter.id")
normal <- as_data_frame(read.table('normalSample', header = F, sep = "\t"))
colnames(normal) <- sampleHeaders
kable(normal, caption = "Normal Sample Information")

tumor <- as_data_frame(read.table('tumorSample', header = T, sep = "\t"))
colnames(tumor) <- sampleHeaders
kable(tumor, caption = "Tumor Sample Information")
```

## Somatic Mutation Calling

1. Download 20 cases of BRCA WXS bam files from GDC-portal.
2. Download Regulatory Build annotation from Ensembl.
3. Call Regulatory element mutations by GATK (DP > 10, AD > 3 )
4. Exclude dbSNP, avsnp147 (ANNOVAR)
5. Find recurrent tumor and nomal respectively.
6. Somatic tumor.recurrent - normal.recurrent.

## Result

```{r load data}
resultHeaders <-c("chrom","pos","ref","alt","ensid","desc","feature","recurrent")
chromOrder = c("chr1","chr2","chr3","chr4","chr5","chr6","chr7","chr8","chr9","chr10","chr11","chr12","chr13","chr14","chr15","chr16","chr17","chr18","chr19","chr20","chr21","chr22","chrX")
somatic <- as_data_frame(read.table('somatic.recurrency.sortByChrom', header = F, sep = "\t"))
colnames(somatic) <- resultHeaders
somatic <- filter(somatic, ensid != "." & recurrent > 1) 
somatic$chrom <- factor(somatic$chrom, levels = chromOrder)
```

### Recurrent count and feature distribution
```{r Recurrent and feature count}
ggplot(somatic, aes(x = factor(recurrent))) + geom_bar(aes(fill = feature)) + geom_text(stat = 'count', aes(label = ..count..), vjust = -0.3) + theme_minimal() + xlab("Recurrent") + ylab("Count") + ggtitle(paste("Total ",nrow(somatic)," (Recurrent > 2)"))
```

### Feature distribution on the chromosome
```{r}
ggplot(somatic, aes(x = chrom)) + geom_bar(aes(fill = feature)) + geom_text(stat = 'count', aes(label = ..count..), vjust = -0.3, size = 3) + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("Chromosome") + ylab("Count") + ggtitle("Chromosome Distribution")
```

### Mutation type
```{r}
ggplot(somatic, aes(ref,alt)) + geom_bin2d() + theme_minimal() + scale_fill_gradient(low = "#00AFBB", high = "#FC3E07")
```

### Filter recurrent larger than or equal to 5
```{r}
somatic5 <- filter(somatic, recurrent >= 5) 
# ggplot(somatic5, aes(x = ensid)) + geom_bar(aes(fill = feature)) + geom_text(stat = 'count', aes(label = ..count..), vjust = -0.3) + theme_minimal() + xlab("Recurrent") + ylab("Count") + ggtitle(paste("Total ",nrow(somatic5)," (Recurrent > 2)")) + theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(somatic5, aes(x = chrom)) + geom_bar(aes(fill = feature)) + geom_text(stat = 'count', aes(label = ..count..), vjust = -0.3, size = 3) + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("Chromosome") + ylab("Count") + ggtitle("Chromosome Distribution")

ensCount <- somatic5 %>% group_by(ensid) %>% summarise(count = n()) %>% arrange(-count)
kable(ensCount)
```

### Annotation from BioMart
```{r}
ensrIDs <- as.character(ensCount$ensid)
FUNCGEN <- useMart("ENSEMBL_MART_FUNCGEN", host = "useast.ensembl.org")

FUNCGEN.dataset <- as_data_frame(listDatasets(FUNCGEN))
FUNCGEN.dataset.hsapiens <- filter(FUNCGEN.dataset, grepl('hsapiens', dataset))
FUNCGEN.dataset.hsapiens

hsapiens.regulatory <- useDataset("hsapiens_regulatory_feature", FUNCGEN)
hsapiens.regulatory.attributes <- listAttributes(hsapiens.regulatory)
hsapiens.regulatory.filters <- listFilters(hsapiens.regulatory)

ensrIDs.regulatory<- getBM(hsapiens.regulatory.attributes$name[-c(10,11)], filters = "regulatory_stable_id", values = ensrIDs, mart = hsapiens.regulatory)

hsapiens.annotation <- useDataset("hsapiens_annotated_feature", FUNCGEN)
hsapiens.annotation.attributes <- listAttributes(hsapiens.annotation)
hsapiens.annotation.filters <- listFilters(hsapiens.annotation)

# ensrIDs.annotation<- getBM(hsapiens.regulatory.attributes$name[-c(10,11)], filters = "regulatory_stable_id", values = ensrIDs, mart = hsapiens.regulatory)

hsapiens.external <- useDataset("hsapiens_external_feature", FUNCGEN)
hsapiens.external.attributes <- listAttributes(hsapiens.external)
hsapiens.external.filters <- listFilters(hsapiens.external)

```

# Done, Find another 20 cases.





