---
title: "Find_target_of_mutation"
author: "C.J. Liu"
date: "12/14/2016"
output:
  html_document:
    depth: 3
    highlight: tango
    number_sections: yes
    theme: united
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    word_document:
      toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r library packages}
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
library(knitr)
library(biomaRt)
```


# Annotate novel mutation

## Ensembl [Regulatory build][1]
Even though, I map mutation to the Ensembl regulatory build release-86 (the same as release-87 except without CTCF bind), the basic statistics Regulatory feature of Ensembl regulatory build release-86.

|feature_type|No.|
|-----|-----|
|Promoter|16384|
|TF binding site|21922|
|Enhancer|36452|
|Promoter Flanking Region|45706|
|Open chromatin|129294|


## Load filtered data
```{r load data}
recur <- read.table("/home/cliu18/liucj/projects/1.Mutation_calling_in_non-condig_region_through_EXOME/5.recalibrateRecurrencyWithCoverage/SNP/refined_goc/somatic5.recur0.1", header = T, sep = "\t")
recur <- as_data_frame(recur)
#arrange(recur, -rate)
```

## Annotation from biomaRt
### Ensembl Genes
Find genes up and down stream(1M bp) of mutation site.
```{r genes}
GENES = useMart("ENSEMBL_MART_ENSEMBL", dataset = "hsapiens_gene_ensembl", host = "useast.ensembl.org")
GENES.ATTRIBUTES <- listAttributes(GENES)
GENES.FILTERS <- listFilters(GENES)

getAroundGenes <- function(x, len){
  
  around.genes<-getBM(attributes = c("ensembl_gene_id","chromosome_name", "start_position", "end_position", "strand","hgnc_symbol","entrezgene", "gene_biotype"), filters = c("chromosome_name", "start", "end", "biotype"), values = list(chromosome_name=x["chrom"], start=as.numeric(x["pos"]) - as.numeric(len), end = as.numeric(x["pos"]) + as.numeric(len), biotype = "protein_coding"), mart = GENES)
  
  # plus strand gene filter nearest start position protein coding gene
  # minus strand gene filter nearest end position protein coding gene
  
  strand.plus <- filter(around.genes, strand == 1 & start_position - as.numeric(x["pos"]) > 0) %>% slice(which.min(start_position - as.numeric(x["pos"]))) %>% mutate(distance = start_position - as.numeric(x["pos"]))
  
  strand.minus <- filter(around.genes, strand == -1 & end_position - as.numeric(x["pos"]) < 0) %>% slice(which.max(end_position - as.numeric(x["pos"]))) %>% mutate(distance = end_position - as.numeric(x["pos"]))
  
   cbind(t(x), rbind(strand.plus, strand.minus))
  # return(result)
  
  # name<- file.path("/home/cliu18/liucj/projects/1.Mutation_calling_in_non-condig_region_through_EXOME/6.mapMutationToGenes/SNP",len,paste(paste("chr", x["chrom"], sep = ""),as.numeric(x["pos"]), x["ensid"], as.numeric(x["recurr"]),len, sep = "_"))
  
  #write.table(around.genes, file = name, quote = F, sep = "\t", row.names = F)
  # return(x)
}

# Find nearest gene to the mutation
nearestGene <- apply(recur, 1, getAroundGenes, len = "100000000")

#apply(recur, 1, getAroundGenes, len = "5000")
#apply(recur, 1, getAroundGenes, len = "500000")
nearestGene.data.frame <-Reduce(function(...) rbind(...), nearestGene)
# test one line to get nearest plus and minus gene.
# test <- head(recur, 1)
# apply(test, 1, getAroundGenes, len = "1000000")

# test.around.genes <- getBM(attributes = c("ensembl_gene_id","chromosome_name", "start_position", "end_position", "strand","hgnc_symbol","entrezgene", "gene_biotype"), filters = c("chromosome_name", "start", "end", "biotype"), values = list(chromosome_name=test$chrom, start=test$pos - 1000000, end = test$pos + 1000000, biotype= "protein_coding"), mart = GENES)

# test.around.genes.plus <- filter(test.around.genes, strand == 1 & start_position - test$pos > 0) %>% slice(which.min(start_position - test$pos)) %>% mutate(distance = start_position - test$pos)

# test.around.genes.minus <- filter(test.around.genes, strand == -1 & end_position - test$pos < 0) %>% slice(which.max(end_position - test$pos)) %>% mutate(distance = end_position - test$pos)

write.table(nearestGene.data.frame, file = "/home/cliu18/liucj/projects/1.Mutation_calling_in_non-condig_region_through_EXOME/6.mapMutationToGenes/SNP/nearestProteinCodingGene.txt", quote = F, sep = "\t", row.names = F)

```

### Ensembl Regulation
```{r regulation}
FUNCGEN <- useMart("ENSEMBL_MART_FUNCGEN", host = "useast.ensembl.org")

FUNCGEN.dataset <- as_data_frame(listDatasets(FUNCGEN))
FUNCGEN.dataset.hsapiens <- filter(FUNCGEN.dataset, grepl('hsapiens', dataset))
FUNCGEN.dataset.hsapiens

hsapiens.regulatory <- useDataset("hsapiens_regulatory_feature", FUNCGEN)
hsapiens.regulatory.attributes <- listAttributes(hsapiens.regulatory)
hsapiens.regulatory.filters <- listFilters(hsapiens.regulatory)

# ensrIDs.regulatory<- getBM(hsapiens.regulatory.attributes$name[-c(10,11)], filters = "regulatory_stable_id", values = ensrIDs, mart = hsapiens.regulatory)

# hsapiens.annotation <- useDataset("hsapiens_annotated_feature", FUNCGEN)
# hsapiens.annotation.attributes <- listAttributes(hsapiens.annotation)
# hsapiens.annotation.filters <- listFilters(hsapiens.annotation)

# ensrIDs.annotation<- getBM(hsapiens.regulatory.attributes$name[-c(10,11)], filters = "regulatory_stable_id", values = ensrIDs, mart = hsapiens.regulatory)

# hsapiens.external <- useDataset("hsapiens_external_feature", FUNCGEN)
# hsapiens.external.attributes <- listAttributes(hsapiens.external)
# hsapiens.external.filters <- listFilters(hsapiens.external)
```











[1]: http://genomebiology.biomedcentral.com/articles/10.1186/s13059-015-0621-5













