---
title: "mutationInPromoterNext"
author: "C.J. Liu"
date: "1/6/2017"
output:
  html_document:
    depth: 3
    highlight: tango
    number_sections: yes
    theme: united
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    word_document:
      toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```

> Plans and working processes are documented in [My Trello][1].

```{r load librares}
library(dplyr)
library(ggplot2)
```

# Check NG paper.
Check those three NG papers, there is one paper deposited all mutations in a website, you can check if our mutation is in their [ftp][2] list.

I download the data to the /liucj/reference/SANGER/cancer/AlexandrovEtAl on the server 1. The file `samples_summary.txt` is malformated. 

## Samples Summary 
```{r Smaples statistics}
samples <- read.table(file = "/home/cliu18/liucj/reference/SANGER/cancer/AlexandrovEtAl/samples_summary.txt", header = T, sep = "\t")

ggplot(data = samples, aes(x = factor(Cancer.Type), fill = Sequencing.Type)) +
  geom_bar() +
  geom_text(stat = 'count', aes(label = ..count..), vjust = 0.6, hjust = 1, angle = 90, size = 3) +
  theme_minimal() + 
  labs(x = "Cancer Type", y = "Counts", title = "Samples Info", fill = "Sequencing Type") +
  scale_x_discrete(position ="bottom") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.text = element_text(size = 8), legend.title = element_text(size = 8))
  
```

## Cancer Type Summary
```{r Cancer Type}
# Load data
cancers <- read.table(file = "/home/cliu18/liucj/reference/SANGER/cancer/AlexandrovEtAl/cancer_type_summary.txt", header = T, sep = "\t")

```


# Analyze KIRC,LUAD
Create analysis file for KIRC and LUAD, and download data to server 2.

In fact, the whole processes of analyzing KIRC and LUAD are the similar to BRCA. 

1. [LUAD data summary](LUAD/analysis/1.TCGA-LUADDataSummary.Rmd), [KIRC data summary](KIRC/analysis/1.TCGA-KIRCDataSummary.Rmd).








[1]:https://trello.com/b/JwU9J02D
[2]:ftp://ftp.sanger.ac.uk/pub/cancer/AlexandrovEtAl/somatic_mutation_data/





